---
title: "filter_join"
author: "MMB"
date: "7/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here) 
library(kableExtra)
```
### Read in data:
```{r  read_data, echo=FALSE}
fish <- read_csv(here("data", "fish.csv"))
kelp_abur <- read_excel(here("data", "kelp.xlsx"), sheet = "abur")
names(fish); summary(fish); head(fish); tail(fish)
names(kelp_abur); summary(kelp_abur); head(kelp_abur); tail(kelp_abur)
```
### dplyr::filter() to conditionally subset by rows
```{r filter1, echo=FALSE}
fish_garibaldi <- fish %>% 
  filter(common_name == "garibaldi")
summary(fish_garibaldi)
fish_mohk <- fish %>% 
  filter(site == "mohk")
summary(fish_mohk)
fish_over50 <- fish %>% 
  filter(total_count >= 50)
summary(fish_over50)
```
## Filter to return rows that match multiple criteria (OR)
```{r filter2, echo=FALSE}
fish_3sp <- fish %>% 
  filter(common_name == "garibaldi" | 
           common_name == "blacksmith" | 
           common_name == "black surfperch")
summary(fish_3sp)
fish_3sp <- fish %>% 
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))
fish_gar_2016 <- fish %>% 
  filter(year == 2016 | common_name == "garibaldi")
summary(fish_gar_2016)
```
## Filter to return rows that match multiple criteria (AND)
```{r filter3, echo=FALSE}
aque_2018 <- fish %>% 
  filter(year == 2018, site == "aque")
aque_2018 <- fish %>% 
  filter(year == 2018 & site == "aque")
aque_2018 <- fish %>% 
  filter(year == 2018) %>% 
  filter(site == "aque")
aque_2018
```
## Filter to return rows that match multiple criteria (AND)
```{r filter4, echo=FALSE}
aque_2018 <- fish %>% 
  filter(year == 2018, site == "aque")
aque_2018 <- fish %>% 
  filter(year == 2018 & site == "aque")
aque_2018 <- fish %>% 
  filter(year == 2018) %>% 
  filter(site == "aque")
aque_2018
# Observations for garibaldi or rock wrasse AND the total_count is less than or equal to 10
low_gb_wr <- fish %>% 
  filter(common_name %in% c("garibaldi", "rock wrasse"), 
         total_count <= 10)
low_gb_wr
```
## stringr::str_detect() to filter by a partial pattern
```{r str_detect, echo=FALSE}
fish_bl <- fish %>% 
  filter(str_detect(common_name, pattern = "black"))
fish_Nit <- fish %>% 
  filter(str_detect(common_name, pattern = "it", negate = TRUE))
```
## dplyr::*_join() to merge data frames (full_join(), left_join(), and inner_join())
```{r join, echo=FALSE}
abur_kelp_fish <- kelp_abur %>% 
  full_join(fish, by = c("year", "site")) 
View(abur_kelp_fish)
kelp_fish_left <- kelp_abur %>% 
  left_join(fish, by = c("year","site"))
kelp_fish_left
kelp_fish_injoin <- kelp_abur %>% 
  inner_join(fish, by = c("year", "site"))
kelp_fish_injoin
```
## filter() and join() in a sequence
```{r join_filter, echo=FALSE }
my_fish_join <- fish %>% 
  filter(year == 2017, site == "abur") %>% 
  left_join(kelp_abur, by = c("year", "site")) %>% 
  mutate(fish_per_frond = total_count / total_fronds)
View(my_fish_join)
```
## An HTML table with kable() and kableExtra
```{r table, echo=FALSE}
kable(my_fish_join)
# kableExtra::kable_styling() to modify HTML table styles
my_fish_join %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = FALSE)
```

