---
title: "my_tidying"
author: "MMB"
date: "6/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Attach the packages we’ll use here
```{r  setup_packages, include=FALSE}
library(tidyverse)
library(janitor)
library(here)
library(readxl)
```
### read_excel() to read in data from an Excel worksheet
```{r}
inverts <- read_excel(here("data", "inverts.xlsx"))
#View(inverts)
names(inverts)
summary(inverts)
```
## reshape from wider-to-longer format
```{r long, echo=FALSE}
inverts_long <- pivot_longer(data = inverts, 
                                    cols = '2016':'2018',
                                    names_to = "year",
                                    values_to = "sp_count")
inverts_long
class(inverts_long$year)
inverts_long <- inverts_long %>% 
  mutate(year = as.numeric(year))
class(inverts_long$year)
```
## convert from longer-to-wider format
```{r wide, echo=FALSE}
inverts_wide <- inverts_long %>% 
  pivot_wider(names_from = common_name, 
                     values_from = sp_count)

inverts_wide
```
## janitor::clean_names() to clean up column names
```{r colnames, echo=FALSE}
inverts_wide <- inverts_wide %>% 
  clean_names()

names(inverts_wide)
```
## tidyr::unite() to merge information from separate columns
```{r unite, echo=FALSE}
inverts_unite <- inverts_long %>% 
  unite(col = "site_year", # What to name the new united column
               c(site, year), # The columns we'll unite (site, year)
               sep = "_") # How to separate the things we're uniting
```
## 
```{r unite2, echo=FALSE}
inverts_moyr <- inverts_long %>% 
  unite(col = "mo_yr", # What to name the new united column
               c(month, year), # The columns we'll unite (site, year)
               sep = "/") 
```
## 
```{r uniteM, echo=FALSE}

inverts_triple_unite <- inverts_long %>% 
  tidyr::unite(col = "year_site_name",
               c(year, site, common_name),
               sep = "-") # Note: this is a dash

head(inverts_triple_unite)
```
## tidyr::separate() to separate information into multiple columns
```{r separate, echo=FALSE}
inverts_sep <- inverts_unite %>% 
  tidyr::separate(site_year, into = c("my_site", "my_year"))
```
## stringr::str_replace() to replace a pattern
```{r replace, echo=FALSE}
ca_abbr <- inverts %>% 
  mutate(
    common_name = 
      str_replace(common_name, 
              pattern = "california", 
              replacement = "CA")
  )
head(inverts)
head(ca_abbr)
```

