---
title: "pivot_lobsters"
author: "MMB"
date: "6/29/2020"
output: html_document
---
## attach libraries
```{r setup, include=TRUE}
library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
```

## Read In Data

```{r input, echo=FALSE}
lobsters <- read_xlsx(here("data/lobsters.xlsx"), skip=4)
head(lobsters)
```
# Explore Data
```{r summary, echo=FALSE}
skimr::skim(lobsters) 
```
# Show Data
```{r, eval=FALSE}
View(lobsters)
```

### In R, we can create the functionality of pivot tables with the same logic: we will tell R to group by something and then summarize by something.
```{r summarize_data, echo=FALSE}
lobsters <- read_xlsx(here("data/lobsters2.xlsx"), skip=4)
lobsters %>%
  group_by(year) %>%
  summarize(count_by_year = n())
lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_siteyear =  n())
siteyear_summary <- lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_siteyear =  n(), 
            mean_size_mm = mean(size_mm, na.rm=TRUE), 
            sd_size_mm = sd(size_mm, na.rm=TRUE),
            median_size_mm = median(size_mm, na.rm = TRUE))
```
## a ggplot option:
```{r plot_grouped}
ggplot(data = siteyear_summary, aes(x = year, y = median_size_mm, color = site)) +
  geom_line() 
ggsave(here("figures", "lobsters-line.png"))
ggplot(siteyear_summary, aes(x = year, y = median_size_mm)) +
  geom_col() +
  facet_wrap(~site)
ggsave(here("figures", "lobsters-col.png"))
```
  
## inspect our new variable 
There are `r nrow(lobsters)` total lobsters included in this report
```{r}
siteyear_summary
```
## make a table with our new variable
```{r}
table_siteyear <- siteyear_summary %>% kable()
table_siteyear
```
## Using Function  dplyr::count()
```{r}

lobsters %>%
  count(site, year)
```
There are `r count(lobsters)` total lobsters included in this summary.
## Adding new columns to Data, Making New Variables
```{r mutation_of_cols}
lobsters_detailed <- lobsters %>%
  mutate(size_m = size_mm / 1000, 
         millenia = 2000,
         observer = "Allison Horst")
```
## Subset Data, Select Variables
```{r subseting_cols}
lobsters_detailed_sub <- lobsters_detailed %>%
  select(date, site, size_m)
```
