---
title: "Creating graphs in R with `ggplot2`"
author: "MMB"
date: "6/29/2020"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R and it's going to be amazing.

```{r setup, echo=FALSE}
library(tidyverse,quietly = T)
library(readxl,quietly = T)
library(here,quietly = T)
library(scales,quietly = T)   # Need the scales package
```
Reading input data from CSV file
```{r read_csv, echo=FALSE}
ca_np <- read_csv(here("data", "ca_np.csv")) #read_xlsx()ORread_xls()
names(ca_np)
head(ca_np)
summary(ca_np)
```
Reading input data from excel file. read_xlsx() or read_xls() could also be used
```{r read_excel, echo=FALSE}
ci_np <- read_excel(here("data", "ci_np.xlsx"))
names(ci_np)
head(ci_np)
summary(ci_np)
```
ggplot graph: Visitors to Channel Islands NP
```{r plot1, echo=TRUE}
gg_base <- ggplot2::ggplot(data = ci_np, aes(x = year, y = visitors)) 
  gg_base +
  geom_area()
  gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
  gg_base + 
  geom_point(color = "purple",
             pch = 17,
             size = 4,
             alpha = 0.5)
  gg_base + 
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  )
  gg_base + 
  geom_col(aes(fill = visitors))
  gg_base +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()
  gg_base +
  geom_line(linetype = "dotted") +
  theme_bw() +
  labs(
    x = "Year",
    y = "Annual park visitors",
    title = "Channel Islands NP Visitation",
    subtitle = "(1963 - 2016)"
  ) + scale_y_continuous(labels=comma) 
  gg_base +
  geom_line(color = "purple") +
  geom_point(color = "orange",
             aes(size = year),
             alpha = 0.5)
  gg_base +
  geom_col(fill = "orange",
           color = "purple") +
  geom_line(color = "green")
  
```
Now plotting multiple series
```{r plot2, echo=TRUE}
gg_np <- ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name))
gg_np + geom_line(aes(color = park_name)) 
gg_np +
  geom_line(show.legend = FALSE) +
  theme_light() + 
  labs(x = "year", y = "annual visitors") +
  facet_wrap(~ park_name)
```
Exporting a ggplot graph with ggsave()
ggsave() will export the most recently run ggplot graph by default (plot = last_plot()), unless you give it the name of a different saved ggplot object. Some common arguments for ggsave():
```{r save_plot}
ggsave(here("figures", "np_graph.jpg"), dpi = 180, width = 8, height = 7)
```

